### 2009-07-17 SC09/conclude.tex

\section{Conclusion}
\label{conclude}

Large systems experience failures.  To protect themselves against these
failures, parallel applications running on these systems save their progress by
checkpointing.  Unfortunately for many of these applications, their preferred
checkpointing patterns are extremely challenging for the \upfs\ and impose
severe bandwidth limitations.  In this paper, we have developed \plfs\ to
demonstrate how a simple interposition layer can transparently rearrange these
challenging patterns and improve checkpoint bandwidth by several orders of
magnitude.

The parallel file system attached to Roadrunner is the largest LANL has ever
had; testing it has revealed that the challenges of N-1 patterns are severely
exacerbated at this scale.  Given current bandwidths, we know of no current N-1
application at LANL that can effectively checkpoint across the full width of
Roadrunner.  \plfs\ allows them to do so.