### 2009-07-17 SC09/metadata-fig.tex

\begin{SCfigure}
    \centering
    \includegraphics[width=0.4\textwidth]{data/stat/stat.eps}
    \mycaption{fig-metadata}{Metadata Queries.}{
        This graph compares the time to \syscall{stat} a 20~\GB\ file
        on \plfs\ with the time to \syscall{stat} a comparable
        file on PanFS.    
        The \yaxis, which is in logscale,
        shows the \syscall{stat} time 
        as a function of the number of writers who created that
        file.  We compare the PanFS \syscall{stat} rate to the
        \plfs\ rate for files opened for writing, and to the 
        \plfs\ rate for closed files.
        \vspace{1cm}
    }
\end{SCfigure}