### 2009-07-17 SC09/summary-fig.tex

\begin{SCfigure}
    \centering
    \includegraphics[width=0.6\textwidth]{data/summary/page1.eps}
    \mycaption{fig-page1}{Summary of our results.}{
        This graph summarizes our results which will be
        explained in detail in Section~\ref{eval}.  
        The key observation here is that our technique
        has improved checkpoint bandwidths for all seven
        studied benchmarks and applications by up to
        several orders of magnitude.  
        \vspace{1.0cm}
    }
\end{SCfigure}