### 2009-07-29 SC09/plfs.tex
got the graphs looking good. got the abstract right. got the keywords, crap, authors, copyright.
git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@7 979fb63a-54b4-4514-8d67-9ab8ea1b59df

%\documentclass[10pt]{article}
\documentclass{sig-alternate}
\pagenumbering{arabic}
\pagestyle{plain}
%\pagestyle{empty}
%\usepackage{fullpage}
%\usepackage{setspace}
%\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
%\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{subfig}
\usepackage{eso-pic,type1cm,color}

\input{macros}
%\input{watermark}

\begin{document}
\conferenceinfo{SC09} {November 14--20, Portland, Oregon, USA.}
\CopyrightYear{2009}
\crdata{978-1-60558-744-8/09/11}
\boilerplate{(c) 2009 Association for Computing Machinery. ACM acknowledges that
this contribution was authored or co-authored by a contractor or
affiliate of the U.S. Government. As such, the Government retains a
nonexclusive, royalty-free right to publish or reproduce this article,
or to allow others to do so, for Government purposes only.}
%\doublespace
\input{title}
\input{abstract}
\category{D.4.3}{Operating Systems}{File Systems Management}{File organization}
\terms{Performance, Design}
\keywords{High performance computing, parallel computing, checkpointing}
\input{intro}
\input{motivation}
\input{arch}
\input{eval}
\input{related}
\input{future}
\input{conclude}
\input{bib}
\end{document}