### 2009-07-29 SC09/future.tex
attached is the most recent version of the paper.  I'll do another big push tomorrow and then I'll submit it since I'll be on vacation all next week.  I think I've addressed most of the reviewer's comments but here's a couple that maybe still need to be addressed (or ignored):
the addition of metadata benchmark such as mdtest would make paper
stronger

the paper does not mention about file locks.  file system semantics is
not clear.

there should be a discussion about file system approach vs library
approach.  library can speed up N-1 w/out sacrificing performance for
other ops like metadata q

\section{Current Status and Future Work}
\label{future}

\scribble{I removed the 'cute' part of this. Sorry, john}
We initially developed \plfs\ in order to test our hypothesis that an
interposition layer could rearrange checkpoint patterns such that the
convenience of N-1 patterns could be preserved while achieving the bandwidth of
N-N.  However, after achieving such large improvements with real LANL
applications, we were compelled to harden it 
into a production file system. Consisting of about three thousand lines of
code, \plfs\ is currently an optional mount point on Roadrunner where several
applications have begun using it to improve their N-1 checkpoint performance.

%Since \plfs\ consists of only three thousand
%source lines of code and is easily ported to all Linux architectures, we
%decided~\footnotemark that the costs of maintaining it are easily outweighed by
%the benefits.
%\footnotetext{were persuaded by management}

Although \plfs\ works transparently with unmodified applications, in practice
some applications may need to make minor changes.  Many users find it most
convenient to stage their input data sets and their executable code in the same
directory where they write their checkpoint files.  Since \plfs\ is
specifically a checkpoint file system and is not a general purpose file system,
this usage pattern may suffer a performance hit.  These users should instead
stage their input data sets and their executable code in a general purpose file
system and merely direct their N-1 checkpoint files to a \plfs\ directory.

Also, applications that open \plfs\ files in read-write mode will find that
reading from these files can be very slow.  As we discussed in
Section~\ref{arch-read}, reads in \plfs\ are handled by reading and aggregating
the index files.  When files are opened in write-read mode, this process must
be repeated for every \syscall{read} since intervening \syscall{writes} may
have occurred.  Although the vast majority of HPC applications are able to read
in read-only mode, we do plan to investigate removing this limitation in the
future for those few applications who must occasionally read while writing.
One possibility is to introduce metadata servers which can synchronously
maintain an aggregated index in memory.

Originally intended merely to address N-1 challenges, \plfs\, as currently
designed, also has the potential to address N-N challenges.  One way that
\plfs\ can reduce N-N checkpointing times is by reducing disk seeks through its
use of log-structured writing.  However, we have yet to measure the
frequency of non-sequential IO within N-N checkpoints.

\scribble{Added gigaplus reference to the end. --Milo}
Another challenge of an N-N pattern is the
overwhelming metadata pressure resulting from the simultaneous creation of tens
of thousands of files within a single directory.  Currently HPC parallel file
systems do distribute metadata across multiple metadata servers; however they do so
at the granularity of a \Term{volume} or a directory (\ie\ all files within a
directory share a single metadata server).  \plfs\ can easily refine this
granularity by distributing container subdirectories 
across multiple metadata servers.  At the moment, \plfs\ only creates container
structures for regular files; directories are simply created directly on the
\upfs.  By extending \plfs\ to create a similar container structure for
directories, we believe that \plfs\ can effectively address this N-N challenge
as well, in a manner similar to~\cite{gigaplus}.