### 2010-04-29 devel_path/table.tex

\begin{table*}[tb]
\centering
\begin{tabular}{|lrrr|}
\hline
TEST & Description Lines & Keys Collected & Transducer Lines \\
\hline
mdtest & 48 & 88 & 200 \\
ior & 45 & 101 & 143 \\
LANL fs\_test & 36 & 149 & 2 \\
\hline
\end{tabular}
\mycaption{eval-table}{Defining Experiments.}{
This table shows how many lines are necessary to describe typical parameter
sweeps for three different applications which are currently using \name\ at
LANL.  It also shows the number of \kv\ pairs collected from each \sub, and the
lines needed for the application-specific transducers which parse the output of
each \sub\ and produce these \kv\ pairs.  The LANL fs\_test transducer is so
small because LANL fs\_test already produces output in \kv\ format so that
transducer simply echoes its input to its output.  The other transducers are
larger because those applications do not have consistent output formatting so
every \kv\ must be parsed differently. 
}
\end{table*}