### 2010-10-05 roadmap/gary-squash.tex

The \plfs\ library needs to support something like this functionality:

\begin{enumerate}
\item{ I have this open fd, please give me the local map for
this fd (this is not for the entire map, it is for my part
only
parms: open fd, pointer that the lib would put an address of some
space the lib mallocs where it puts the map part, and a length of
the map part}
\item{ please store this squashed entire map in this
specified subcontainer
parms: fd, address of map, length of map, container to store}
\item{ mark the metadata for this fd as successfully squashed and distributed
this tells the lib to put a metadata dropping in
parms: fd}

\item{ using this consolidated set of map blobs , make me a squashed map
parms: addr of consolidated map blobs, length of consolidatd map blobs,
addr to put squashed map, length of resultant squashed map}

\item{ is there a metadata dropping that this has been squashed and
distributed for this file and where are the locations one might
look for a replica - this gives a valid list of places where
indexes live and whether or not there is a valid squashed index
(subcontainers),
may need an fd and file path version of this or a flag for lookup type
parms: file path, fd, addr for lib to put location list, length of list}

\item{ open for read with squashed index replica location or a squashed map
provided in memory
parms: open stuff plus replica location or map squashed provided}
\end{enumerate}

So an mpi-io adio app on would do "something like" this:
\\
open for write
\\
write stuff
\\
on close for write "this close for write" is a bit of a complication"
(I include frees required by the lib - other malloc/free would
be required by the adio layer itself)
- all would ask for their fd (local) index blob  (new func 1)
 (I suppose we didnt write anything is a possible outcome)
\\
- all reduce these local index blobs to proc 0
\\
- proc 0 calls using these parts make me a squashed map (new func 4)
\\
- proc 0 calls is there squashed metadata for this file
  which gives proc 0 a list of subcontainers (new func 5)
\\
- proc 0 assigns work (squashed map and subcontainer) to get
  all the squashed maps distributed
\\
- all workers call please store this squashed map in this subcontainer
(new func 2)
\\
- all free the blobs local blobs (allocated by the lib)
\\
- proc 0 frees squashed map (allocated by the lib)
\\
- all but proc 0 barrier
\\
- proc 0 mark metadata for this fd as successfully squashed and
replicated  (new func 3)
\\
- proc 0 barrier
\\
- all call close

So you can see that the lib doesnt do any mpi magic and the adio (or
app) doesnt know what is in the map or a subcontainer, they are just
blobs of data used to distribute, because it has the distribution
mechanism.

You notice that the distribution of the squashed indexes to
the brethren does not just blindly do a broadcast and write
the index, because you might want to write fewer indexes
than procs.

So this gets the index squashed on close for write by adio.
We might want a hint on whether you want squashed indexes
on this mpi run so you could avoid this if you dont want/need
it.  Probably should plan on that up front even if we dont
want to use it.

Now the adio on open for read would
proc 0 in the adio layer would check to see if there
is there a valid squashed index and get the location list  (new func 5)
\\
--if valid squashed index
\\
----  assign replicas to the brethren
\\
---- everyone opens using that squashed index  (few func 6 using
subcontainer)
\\
--if there is no valid squashed index
\\
---- proc 0 has a list of places where indexes live - from the query for
squash above - even though there is no squashed index,
\\
---- proc zero assigns work to the brethren (may not be equal number of
workers and number of indexes to read),
\\
---- each proc does open read with their index part assigned (new func 6
using subcontainer)
\\
---- get local map (new func 1)
\\
---- send it back to proc 0,
\\
---- when proc 0 gets them all it calls build me a squashed index from
these parts into the lib (new func 4)
\\
---- broadcast resultant squashed map to all,
\\
---- everyone opens with squashed map provided (new func 6 - using mem
map) (I suppose you could write the squashed maps to disk here but that
seems bad on a read op)

adio close on read has nothing much to do maybe frees or something,
again you could write the squashed indexes here but on a read
that is probably bad

A stand alone mpi utility to squash would then:
\\
- just be an mpi program that open rw (because it is open for read it
would efficiently get the squashed map with as many procs as you have
available - doesnt have to match the number of containers) - everyone
would end up with a copy of the fully squashed map
\\
- close for rw  (Because this is a close for write the squashed indexes
are written)

A stand alone non mpi utility to squash would be essentially the same
thing as what the mpi open for rw and close for write is, but
the farm out the work would be a loop to do the work and not to assign
the work, or maybe do it with threads where there was procs, which
would make the stand alone utility to be basically the same as the mpi
utility.

Again, this close for read/close for write seems to be a slight
complication, but maybe an important capability.

Of course this is all notional, but

\begin{enumerate}
\item{this seems doable}
\item{works where num subcontainers doesnt equal num procs}
\item{leverages mpi flie open and close}
\item{gets us easy stand alone mpi and non mpi squasher}
\item{nearly optimal for mpi and helps posix hugely}
\end{enumerate}

Now the posix open for read does something similar to
what proc zero and the other procs do, it would need
to call the same sorts of routines.  It would probably
just need to randomly pick a subcontainer to
do the open with container for squashed index.
The posix close probably wouldnt do a squash, as it
wont know when people are done and how to do it and
who else is doing it etc.