### 2012-09-17 hdf5/discs_2012/abstract.tex
Adding existing files

HDF5 is a data model, library and file format for storing and managing data. It is designed for flexible and efficient I/O for high volume and complex data. Natively, it uses a single-file format where multiple HDF5 objects are stored in a single file. In a parallel HDF5 application, multiple processes access a single file, thereby resulting in a performance bottleneck in I/O. Additionally, a single-file format does not allow semantic post processing on individual objects outside the scope of the HDF5 application. The HDF group has recently introduced a new layer called the Virtual Object Layer (VOL) that allows writing plugins for HDF5 to store data in different ways. We have written a new plugin for HDF5 using PLFS that provides serves two purposes: 1) it uses PLFS to convert the single-file layout into a data layout that is optimized for the underlying file system, and 2) it stores data in a unique way that enables semantic post-processing on the data, independently of HDF5. In this paper, we measure the performance improvements available through the PLFS VOL and discuss future work leveraging the new semantic post-processing functionality enabled.  We further discuss the applicability of this approach for exascale burst buffer storage systems.